---
title: "Introduction to frameR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to frameR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE
)
```

## Overview

`frameR` measures emphasis frames in political text using
transformer-based sentence embeddings and cosine similarity.
Rather than classifying sentences into frame categories — which
measures how often a frame appears — `frameR` measures the semantic
association between issue-relevant discourse and theoretically
specified evaluative dimensions. This operationalizes emphasis frames
as they are conceptualized in the political communication literature:
as semantic links between a core issue and peripheral considerations
that shape how that issue is perceived and evaluated (Druckman 2011).

The package is designed for researchers who want to:

- Track how a party's framing of an issue evolves across time
- Compare framing strategies across parties on a shared issue
- Conduct cross-national comparisons of framing patterns across
  languages and party families

All code chunks in this vignette use `eval = FALSE` because the
pipeline requires a Python environment set up separately via
`install_frameR()`. See the Installation section below.

## Installation

Install `frameR` from GitHub:

```{r install}
devtools::install_github("spenccorp/frameR")
```

After installing, set up the Python environment once and restart R:

```{r setup_python}
library(frameR)
install_frameR()
# Restart R when prompted, then load the package again
library(frameR)
```

Check the environment is ready:

```{r status}
frameR_status()
```

## A Complete Example: Immigration Framing Across French Parties

This example demonstrates the full `frameR` pipeline using real
manifesto data. We compare how the French National Rally and the
Socialist Party framed immigration in their 2022 electoral manifestos,
using quasi-sentences from the Comparative Manifesto Project as
collected and distributed with this package.

This illustrates a cross-party comparison research design: do
parties with different ideological positions frame the same issue
in semantically distinct ways? Frame theory predicts yes — not only
should their policy positions differ, but the evaluative dimensions
through which they discuss immigration should differ as well.

### The Data

`frameR` includes a built-in dataset of 4,240 quasi-sentences from
the 2022 French legislative election manifestos of the National
Rally (RN, party code 31720) and the Socialist Party (PS, party
code 31320), as collected by the Comparative Manifesto Project.

```{r data}
library(frameR)

# Load the bundled example data
data(vignette_data)

# Inspect
head(vignette_data)
nrow(vignette_data)
table(vignette_data$party_name)
```

### Step 1: Load a Model

Load the MARPOR-adapted embedding model. This model has been
fine-tuned on EU27 party manifestos and is the recommended default
for political text applications.

```{r load_model}
model <- load_model()
```

For applications outside the EU27 manifesto context, the base
multilingual model is available:

```{r base_model}
model <- load_model("base")
```

### Step 2: Embed the Corpus

Embed all sentences once. Embeddings are stored and reused across
all downstream steps.

```{r embed}
embeddings <- embed_corpus(vignette_data$text, model)
dim(embeddings) # 4240 x 768
```

### Step 3: Identify Immigration-Relevant Sentences

Identify sentences discussing immigration using the dual-criteria
approach: keyword matching combined with semantic similarity
thresholding. Keywords are specified in French to match the corpus
language.

```{r identify}
topic_mask <- identify_topic_sentences(
  sentences    = vignette_data$text,
  embeddings   = embeddings,
  anchor_words = c(
    "immigration", "immigré", "immigrés",
    "migration", "migrant", "migrants",
    "asylum", "réfugié", "réfugiés"
  ),
  model        = model,
  threshold    = 0.5,
  method       = "both"
)

# How many immigration sentences per party?
table(vignette_data$party_name[topic_mask])
```

We recommend checking robustness to the threshold value by running
the identification step under alternative thresholds (0.3, 0.5, 0.7).

### Step 4: Specify and Embed Frame Categories

Specify frame categories derived from the theoretical literature on
far-right immigration framing (Helbling 2014; Rydgren 2008). Keywords
are specified in French.

```{r frames}
frame_categories <- list(
  Economic = c(
    "économie", "emplois", "emploi",
    "welfare", "État providence", "coût", "charges"
  ),
  Cultural = c(
    "culture", "valeurs", "tradition",
    "identité", "islam", "laïcité", "nation"
  ),
  Security = c(
    "crime", "frontière", "terrorisme",
    "sécurité", "illégal", "violence"
  )
)

keyword_embeddings <- embed_keywords(frame_categories, model)
```

### Step 5: Measure Frame Strength

Measure frame strength for each party treating party membership
as the comparison dimension. We use party codes as period labels
since `measure_frames` compares exactly two groups.

```{r measure}
# Assign synthetic years by party for comparison
# RN = period 1, PS = period 2
comparison_years <- ifelse(
  vignette_data$party_name == "National Rally", 1, 2
)

results <- measure_frames(
  embeddings         = embeddings,
  topic_mask         = topic_mask,
  keyword_embeddings = keyword_embeddings,
  periods            = list(
    `National Rally`   = c(1, 1),
    `Socialist Party`  = c(2, 2)
  ),
  years              = comparison_years,
  n_bootstrap        = 10000L,
  n_permutations     = 10000L,
  seed               = 42L
)
```

### Step 6: Inspect and Plot Results

```{r print}
print(results)
```

```{r plot}
plot(results)
```

## Interpreting Results

### Absolute frame similarity

Higher cosine similarity indicates stronger semantic association
between a party's immigration discourse and a frame category's
evaluative considerations. These scores are not interpretable in
absolute terms — what matters is the comparison between parties.

### Relative frame emphasis

Relative emphasis scores normalize similarities within each party
so they sum to one. These reveal which frame dominates each party's
immigration discourse regardless of absolute similarity levels.

### Between-group differences

The permutation test assesses whether observed differences between
parties are statistically distinguishable from chance. A positive
difference indicates the Socialist Party scores higher on that frame;
a negative difference indicates the National Rally scores higher.

## Cross-Lingual Applications

`frameR` supports cross-lingual analysis through the multilingual
MARPOR-adapted model. The same pipeline applies to corpora in any
of the 24 EU official languages — simply specify anchor words and
frame keywords in the language of your target corpus.

```{r multilingual}
# German immigration keywords
anchor_de <- c("Einwanderung", "Einwanderer", "Migration",
               "Asyl", "Flüchtling")

frames_de <- list(
  Economic = c("Wirtschaft", "Arbeitsplätze", "Sozialsystem",
               "Kosten", "Löhne"),
  Cultural = c("Kultur", "Werte", "Tradition",
               "Identität", "Islam", "Nation"),
  Security = c("Kriminalität", "Grenze", "Terrorismus",
               "Sicherheit", "illegal")
)
```

## Domain Adaptation

The default `frameR-marpor` model has been fine-tuned on EU27
party manifestos from the Comparative Manifesto Project, covering
24 languages and decades of electoral competition. This adaptation
makes the embedding space more sensitive to the semantic register
of political manifesto text.

For applications outside this domain, fine-tune a custom model:

```{r custom_model}
# Fine-tune on your own domain corpus
# (exclude your target corpus to avoid circularity)
custom_model <- load_model("base")
# See ?load_model for fine-tuning utilities
```

## References

Druckman, James N. (2011). "What's it all about? Framing in
political science." In *Perspectives on framing*, pp. 282-296.
Psychology Press.

Helbling, Marc (2014). "Framing immigration in western Europe."
*Journal of Ethnic and Migration Studies* 40(1): 21-41.

Reimers, Nils and Iryna Gurevych (2019). "Sentence-BERT: Sentence
embeddings using Siamese BERT-networks." In *Proceedings of EMNLP*.

Rydgren, Jens (2008). "Immigration sceptics, xenophobes or racists?
Radical right-wing voting in six West European countries."
*European Journal of Political Research* 47(6): 737-765.

Volkens, Andrea, et al. (2021). The Manifesto Data Collection.
Manifesto Project (MRG/CMP/MARPOR). Version 2021b. Berlin:
Wissenschaftszentrum Berlin für Sozialforschung (WZB).
