---
title: "Introduction to frameR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to frameR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  eval      = FALSE
)
```


## Overview

`frameR` measures emphasis frames in political text using 
transformer-based sentence embeddings and cosine similarity. 
Rather than classifying sentences into frame categories — which 
measures how often a frame appears — `frameR` measures the semantic 
association between issue-relevant discourse and theoretically 
specified evaluative dimensions. This operationalizes emphasis frames 
as they are conceptualized in the political communication literature: 
as semantic links between a core issue and peripheral considerations 
that shape how that issue is perceived and evaluated (Druckman 2011).

The package is designed for researchers who want to:

- Track how a party's framing of an issue evolves across time
- Compare framing strategies across parties on a shared issue
- Conduct cross-national comparisons of framing patterns across 
  languages and party families

All examples in this vignette use `eval = FALSE` because the pipeline
requires a Python environment that is set up separately. See the 
Installation section for setup instructions.

## Installation

Install `frameR` from GitHub:

```{r install, eval = FALSE}
devtools::install_github("spenccorp/frameR")
```

After installing, set up the Python environment once and restart R:

```{r setup_python, eval = FALSE}
library(frameR)
install_frameR()
# Restart R when prompted, then load the package again
library(frameR)
```

Check the environment is ready:

```{r status, eval = FALSE}
frameR_status()
```

## A Complete Example: Immigration Framing in French Party Manifestos

This example walks through the full `frameR` pipeline using a 
simplified version of the analysis presented in the accompanying 
methods paper. We examine whether the French National Rally shifted 
its immigration framing between an early period (1997-2011) and a 
later period (2012-2022).

### Step 1: Load a Model

Load the base sentence embedding model. For political manifesto 
text we recommend the MARPOR-adapted model when available — see
`?load_model` for details.

```{r load_model}
library(frameR)

model <- load_model("base")
```

### Step 2: Prepare Your Corpus

`frameR` expects a character vector of sentences and a corresponding 
numeric vector of years. Your corpus should be pre-processed into 
individual sentences before passing to the package.

```{r prepare_corpus}
# In practice, load your corpus from a CSV or data frame:
# corpus <- read.csv("my_manifesto_corpus.csv")
# sentences <- corpus$text
# years     <- corpus$year

# For illustration we use a small example corpus
sentences <- c(
  "L'immigration coûte cher à l'État providence.",
  "Les immigrés prennent les emplois des travailleurs français.",
  "Notre culture et nos valeurs sont menacées.",
  "Nous devons protéger l'identité nationale.",
  "La sécurité aux frontières est essentielle.",
  "La criminalité augmente avec l'immigration illégale.",
  "L'économie souffre d'une migration incontrôlée.",
  "La France ne peut plus se permettre l'immigration.",
  "L'immigration massive remet en cause nos valeurs républicaines.",
  "Il faut protéger les femmes contre les cultures incompatibles."
)

years <- c(
  1997, 1997, 2002, 2002,
  2012, 2012, 2017, 2017,
  2022, 2022
)
```

### Step 3: Embed the Corpus

Embed all sentences using the loaded model. Embeddings are computed
once and reused across all downstream steps — do not re-embed between
steps.

```{r embed}
embeddings <- embed_corpus(sentences, model)

# embeddings is a matrix: one row per sentence, 768 columns
dim(embeddings)
```

### Step 4: Identify Issue-Relevant Sentences

Identify which sentences discuss immigration using the recommended
dual-criteria approach: keyword matching combined with semantic
similarity thresholding.

```{r identify}
topic_mask <- identify_topic_sentences(
  sentences    = sentences,
  embeddings   = embeddings,
  anchor_words = c("immigration", "immigrés", 
                   "migration", "immigré"),
  model        = model,
  threshold    = 0.5,
  method       = "both"
)

# Check how many sentences were identified
sum(topic_mask)

# Inspect identified sentences
sentences[topic_mask]
```

We recommend checking robustness to the threshold value by running
the identification step under alternative thresholds (0.3, 0.5, 0.7)
and confirming that substantive findings are stable.

### Step 5: Specify and Embed Frame Categories

Specify frame categories as named lists of keywords, derived from
the theoretical and empirical literature on your issue and context.
Keywords should be in the language of your target corpus.

```{r frames}
frame_categories <- list(
  Economic = c(
    "économie", "emplois", "emploi", 
    "welfare", "État providence"
  ),
  Cultural = c(
    "culture", "valeurs", "tradition", 
    "égalité", "islam", "laïcité"
  ),
  Security = c(
    "crime", "frontière", "terrorisme", 
    "sécurité", "illégal"
  )
)

keyword_embeddings <- embed_keywords(frame_categories, model)
```

#### Robustness to Keyword Specification

Because frame categories are researcher-specified, we recommend
checking robustness under broader and stricter keyword sets. Here
is an example for the Economic frame:

```{r robustness}
# Broader Economic specification
frames_broader <- list(
  Economic = c(
    "économie", "emplois", "emploi", "welfare",
    "État providence", "travail", "salaire", 
    "chômage", "coût", "charges"
  ),
  Cultural = c(
    "culture", "valeurs", "tradition", "égalité",
    "islam", "laïcité", "identité", "nation",
    "patrimoine", "intégration"
  ),
  Security = c(
    "crime", "frontière", "terrorisme", "sécurité",
    "illégal", "police", "délinquance", "violence",
    "ordre", "contrôle"
  )
)

# Stricter Economic specification  
frames_stricter <- list(
  Economic = c("économie", "emploi", "chômage"),
  Cultural = c("culture", "tradition", "laïcité"),
  Security = c("terrorisme", "sécurité", "frontière")
)
```

Run the full analysis under each specification and confirm that
directional findings are consistent.

### Step 6: Measure Frame Strength

Measure frame strength across time periods using bootstrap inference
and permutation tests.

```{r measure}
results <- measure_frames(
  embeddings         = embeddings,
  topic_mask         = topic_mask,
  keyword_embeddings = keyword_embeddings,
  periods            = list(
    Latency    = c(1997, 2002),
    Activation = c(2012, 2022)
  ),
  years              = years,
  n_bootstrap        = 10000L,
  n_permutations     = 10000L,
  seed               = 42L
)
```

### Step 7: Inspect and Plot Results

```{r print}
print(results)
```

The print output shows:

- **Frame Similarity**: Bootstrap mean and 95% confidence interval 
  for each frame category and period. Higher values indicate stronger
  semantic association between issue discourse and that frame category.
- **Between-Period Differences**: The change in mean frame similarity
  from the first to the second period, with permutation test p-values.
  Positive values indicate increased framing in the second period.
- **Relative Frame Emphasis**: Frame similarities normalized within
  each period so they sum to 1. Use these to assess which frame
  is dominant within a period, not to compare absolute levels across
  frame categories.

```{r plot}
# Full three-panel plot
plot(results)

# Individual panels
plot(results, type = "similarity")
plot(results, type = "ratios")
plot(results, type = "differences")
```

## Interpreting Results

### What cosine similarity values mean

Cosine similarity scores are not interpretable in absolute terms.
A similarity of 0.4 to the Economic frame keywords does not mean 
the party uses economic frames 40% of the time. What is meaningful
is comparison: an increase from 0.3 to 0.4 across periods indicates
that the discourse has become more semantically proximate to economic
considerations.

### Comparing across frame categories

Do not compare absolute similarity scores across frame categories
within a single period. Some concepts are inherently closer together
in embedding space than others for reasons that reflect general 
language patterns rather than party strategy. Use the relative
emphasis (ratio) scores for within-period cross-frame comparison.

### Statistical significance

P-values from the permutation test indicate whether between-period
differences are distinguishable from chance. Report both the magnitude
of the difference and the p-value — statistical significance alone
carries limited meaning when corpora are very large or effect sizes
are very small.

## Cross-Lingual Applications

`frameR` supports cross-lingual analysis through multilingual
embedding models. For cross-national comparisons spanning multiple
languages, use a multilingual model:

```{r multilingual}
# Load a multilingual model
model_multi <- load_model(
  "sentence-transformers/paraphrase-multilingual-mpnet-base-v2"
)
```

When applying the method across languages, translate and validate
frame category keywords independently for each language rather than
translating mechanically. Frame-associated vocabulary does not always
translate directly across political and cultural contexts.

## Domain Adaptation

For best results with political manifesto text, fine-tune the 
embedding model on a domain-relevant corpus before analysis.
Fine-tuning makes the embedding space more sensitive to the semantic
register of political text, producing more precise frame measures.

The fine-tuning corpus should be:

- From the same genre as your target corpus (e.g. party manifestos)
- Distinct from your target corpus to avoid circularity
- Reasonably large (several hundred documents minimum)

A model pre-trained on the full Comparative Manifesto Project corpus
will be made available on Hugging Face upon publication of the
accompanying methods paper. Once available it can be loaded as the
default:

```{r marpor_model, eval = FALSE}
# Coming soon - MARPOR-adapted model
model <- load_model("spenccorp/frameR-marpor")
```

## References

Druckman, James N. (2011). "What's it all about? Framing in 
political science." In *Perspectives on framing*, pp. 282-296.
Psychology Press.

Reimers, Nils and Iryna Gurevych (2019). "Sentence-BERT: Sentence 
embeddings using Siamese BERT-networks." In *Proceedings of EMNLP*.
